Lab 11 - Missing Data and Imputation:Answers

### 2023-02-15 Lab 11 Missing data/Lab11_MissingData.Rmd
Add files via upload

---
title: 'Lab 11 - Missing Data and Imputation'
author: "Eric Parajon"
header-includes:
  - \usepackage{enumerate}
  - \usepackage{graphicx}
output: pdf_document
---

@@

#Using texreg package to display results in a table
texreg(pool(model_mice))

@@

```{r, message = FALSE}

#Load in packages
pacman::p_load(Amelia,
               lubridate)

@@

#Pooling models (using mice)
summary(pool(models_amelia))

#Putting into table
texreg(pool(models_amelia))

@@

Compare how models from mice, Amelia, and the above methods perform to our true data:
```{r, eval = FALSE}
#true model
summary(model_true)

#listwise
summary(model_listwise)

#recoding
summary(model_recode)

#mice
summary(pool(model_mice))

#Amelia
summary(pool(models_amelia))
```

**On your own:**

@@

#use Amelia to generate 5 imputed datasets
#use ts and cs so imputation algerithem knows which variables represent countries and years

@@

#report results, including standard errors

>>>>

---
title: "Lab 11 - Missing Data and Imputation"
author: "Eric Parajon"
header-includes:
- \usepackage{enumerate}
- \usepackage{graphicx}
output:
  html_document:
    df_print: paged
---

@@

#Using htmlreg package to display results in a table
htmlreg(pool(model_mice))

@@

```{r, message = FALSE, results='asis'}

#Load in packages
pacman::p_load(Amelia,
               lubridate,
               mice,
               texreg)

@@

summary(pool(models_amelia))

#Putting into table
htmlreg(pool(models_amelia))

@@

Compare how models from mice, Amelia, and the above methods perform to our true data:
```{r, eval = FALSE, results='asis'}
# Estimate models
mod <- list()
mod[['True']] <- model_true
mod[['Listwise deletion']] <- model_listwise
mod[['Recoding']] <- model_recode
mod[['Mice']] <- pool(model_mice)
mod[['Amelia']] <- pool(models_amelia)

htmlreg(mod)

# Exercise **On your own:**

@@

#use Amelia to generate 5 imputed datasets
#use ts and cs so imputation algorithm knows which variables represent countries and years

@@

#report results, including standard errors
htmlreg(pool(___))

#or if knitting to pdf
#texreg(pool(___))

### 2023-02-15 Lab 11 Missing data/Lab11_MissingData_AnswerKey.Rmd
Add files via upload

---
title: 'Lab 11 - Missing Data and Imputation:Answers'
author: "Eric Parajon"
header-includes:
  - \usepackage{enumerate}
  - \usepackage{graphicx}
output: pdf_document
---

@@

**Data!**
First, let's get our data set up. 
```{r}
#clear global environ
rm(list=ls())

@@

```{r, eval = FALSE}
#Load in packages
pacman::p_load(mice,
               VIM,
               texreg) #Visualization

@@

#vis using aggr() from VIM
aggr_plot <- aggr(data_miss, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(data_miss), cex.axis=.7, gap=3, ylab=c("Histogram of missing data","Pattern"))
#what is this showing us?

#MICE TIME
mice_data <- mice(data_miss, #call our data
                  m=5, #number of multiple imputations
                  maxit=50, #number of iterations
                  # method = ? mice will decide the most appropriate method based on your data
                  seed=500)

@@

So this is great, but then what do we do if we want to run a model? We have five data sets. One option is to pool the data this way:
```{r}
#pooling
model_mice <- with(mice_data,lm(Temp~ Ozone + Solar.R + Wind))
summary(pool(model_mice))

#Using texreg package to display
texreg(pool(model_mice))
```
So mice is great. Let's also learn about Amelia II (my personal favorite, named after Amelia Earhart, because she is missing). Amelia is a little different because it assumes that your data is jointly distributed as multivariate normal. It uses the "expectation-maximization with bootstrapping" as an algorithm for imputation. The EM algorithm alternates between something called the expectation (E-step) and maximization (M-step) steps until it converges on values that were missing - it converges when the current and previous values are quite close (sounds a bit like Newton Raphson, no?). It brings in bootstraping by doing this process on multiple bootstrapped samples drawn from your original (and incomplete) data.

@@

```{r}
#Load in packages
pacman::p_load(Amelia,
               lubridate)

@@

#Putting into table
texreg(pool(models_amelia))

@@

Compare how models from mice, Amelia, and the above methods perform to our true data:
```{r}

@@

texreg(mod)

@@

#Exercise **On your own:**

@@

```{r}
library(Amelia)
data("africa") #read in

@@

#use Amelia to generate 5 imputed datasets
#use ts and cs so imputation algs knows countries and years

@@

#Putting into table
texreg(pool(models_africa))

>>>>

---
title: "Lab 11 - Missing Data and Imputation:Answers"
author: "Eric Parajon"
header-includes:
- \usepackage{enumerate}
- \usepackage{graphicx}
output:
  html_document:
    df_print: paged
---

@@

**Data!**
First, let's get our data set up. 
```{r}
#clear global environment
rm(list=ls())

@@

```{r, eval = TRUE, results='asis'}
#Load in packages
pacman::p_load(mice,
               VIM,
               texreg) #Visualization

@@

#vis using aggr() from VIM
aggr_plot <- aggr(data_miss, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(data_miss), cex.axis=.7, gap=3, ylab=c("Histogram of missing data","Pattern"))
#what is this showing us?
```

```{r, eval = TRUE, results='hide'}
#MICE TIME
mice_data <- mice::mice(data_miss, #call our data
                  m=5, #number of multiple imputations
                  maxit=50, #number of iterations
                  seed=500)

@@

So this is great, but then what do we do if we want to run a model? We have five data sets. One option is to pool the data this way:
```{r, eval = TRUE, results='asis'}
#pooling
model_mice <- with(mice_data,lm(Temp~ Ozone + Solar.R + Wind))

summary(pool(model_mice))

#Using texreg package to display
htmlreg(pool(model_mice))
```
So mice is great. Let's also learn about Amelia II (my personal favorite, named after Amelia Earhart, because she is missing). Amelia is a little different because it assumes that your data is jointly distributed as multivariate normal. It uses the "expectation-maximization with bootstrapping" as an algorithm for imputation. The EM algorithm alternates between something called the expectation (E-step) and maximization (M-step) steps until it converges on values that were missing - it converges when the current and previous values are quite close (sounds a bit like Newton Raphson, no?). It brings in bootstraping by doing this process on multiple bootstrapped samples drawn from your original (and incomplete) data.

@@

```{r, eval = TRUE, results='asis'}
#Load in packages
pacman::p_load(Amelia,
               lubridate)

@@

#Putting into table
htmlreg(pool(models_amelia))

@@

Compare how models from mice, Amelia, and the above methods perform to our true data:
```{r, eval = TRUE, results='asis'}

@@

htmlreg(mod)

@@

# Exercise **On your own:**

@@

```{r, eval = TRUE, results='asis'}
library(Amelia)
data("africa") #read in

@@

#use Amelia to generate 5 imputed datasets
#use ts and cs so imputation algorithm also knows countries and years

@@

#Putting into table

coefficient.names <-c("Intercept","Inflation","Trade","Civil Liberty","Population")

htmlreg(pool(models_africa),custom.coef.names = coefficient.names)
